SunPy - Python for Solar Physics

### 2014-01-28 2-1-Map.tex

In addition to the data-type classes, the \texttt{map} sub-package provides two 
collection classes, \texttt{CompositeMap} and \texttt{MapCube}, for 
temporally and spatially aligned data respectively.
\texttt{CompositeMap} provides methods for overlaying spatially aligned 
data, with support for visualisation of images and contour lines overlaid 
upon each other.
\texttt{MapCube} 
provides methods for animation of its series of \texttt{Map} objects. 
Listings~\ref{code:compmap_1} and \ref{code:mapcube_1} show how to interact 
with these classes.
In addition, Listing~\ref{code:mapcube_2} shows how the \texttt{MapCube} 
classes interacts with the \texttt{matplotlib.animation} submodule to save a 
video.

>>>>

In addition to the data-type classes, the \texttt{map} sub-package provides two 
collection classes, \texttt{CompositeMap} and \texttt{MapCube}, for 
temporally and spatially aligned data respectively.
\texttt{CompositeMap} provides methods for overlaying spatially aligned 
data, with support for visualisation of images and contour lines overlaid 
upon each other.
\texttt{MapCube} 
provides methods for animation of its series of \texttt{Map} objects. 
Listings~\ref{code:compmap_1} and \ref{code:mapcube_1} show how to interact 
with these classes.
In addition, Listing~\ref{code:mapcube_2} shows how the \texttt{MapCube} 
classes interact with the \texttt{matplotlib.animation} submodule to save a 
video.

### 2014-01-28 2-2-Lightcurve.tex
tiny edits to Map and Lightcurve

Currently, the \texttt{LightCurve} class is compatible with the
following data sources: the GOES X-ray Sensor (XRS), PROBA2/LYRA, and
the SDO EUV Variability Experiment (EVE; only the level ``OCS'' and
% Do we need to explain what OCS means?
average CSV files -- see \url{http://lasp.colorado.edu/home/eve/data/}
for more detail).  For each of these instruments, a subclass of the
\texttt{LightCurve} object is initialised
(\textit{e.g.}, \texttt{GOESLightCurve}) which inherits from
\texttt{LightCurve} but allows instrument-specific functionality to be
included.  Future developments will introduce support for additional
instruments and data products, as well as implementing a factory interface 
similar to that of \texttt{Map}.  Since there is no established standard
as to how time-series data should be stored and distributed, each SunPy 
\texttt{LightCurve} object provides the ability to download its corresponding 
specific data format in its constructor and parse that file type.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import lightcurve
>>> from sunpy.time import TimeRange
>>> goes = lightcurve.GOESLightCurve.create('2011-06-07 06:00',
...                                         '2011-06-07 10:00')
>>> goes.peek()
>>> resampled_goes = goes.data.resample("1H")
% Does 1h resample makes sense?? I vote for an example with some utility
% an example is an example right.
\end{minted}
\begin{center}
\includegraphics[width=10cm]{goes_lightcurve.pdf}
\end{center}
\caption{Example retrieval of a GOES lightcurve for the time interval 
06:00--08:00 UT on 2011 June 7 using a time range, and the output of the 
\texttt{peek()} method.
Finally the data is resampled to a 1 hour cadence.}
\label{code:goes_lc}
\end{listing}
%schriste - note to self, fix this example, add resampled points on top of plot?

@@

%Stuart: Where did the pandas section go?

>>>>

Currently, the \texttt{LightCurve} class is compatible with the
following data sources: the GOES X-ray Sensor (XRS), PROBA2/LYRA, and
the SDO EUV Variability Experiment (EVE; only the level ``OCS'' and
% Do we need to explain what OCS means?
average CSV files -- see \url{http://lasp.colorado.edu/home/eve/data/}
for more detail).  For each of these instruments, a subclass of the
\texttt{LightCurve} object is initialised
(\textit{e.g.}, \texttt{GOESLightCurve}) which inherits from
\texttt{LightCurve} but allows instrument-specific functionality to be
included.  Future developments will introduce support for additional
instruments and data products, as well as implementing a factory interface 
similar to that of \texttt{Map}.  Since there is no established standard
as to how time-series data should be stored and distributed, each SunPy 
\texttt{LightCurve} object sub-class provides the ability to download its corresponding 
specific data format in its constructor and parse that file type.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import lightcurve
>>> from sunpy.time import TimeRange
>>> goes = lightcurve.GOESLightCurve.create('2011-06-07 06:00',
...                                         '2011-06-07 08:00')
>>> goes.peek()
%removed the resample part of the example - really we need to fix truncate!
\end{minted}
\begin{center}
\includegraphics[width=10cm]{goes_lightcurve.pdf}
\end{center}
\caption{Example retrieval of a GOES lightcurve for the time interval 
06:00--08:00 UT on 2011 June 7 using a time range, and the output of the 
\texttt{peek()} method.}
\label{code:goes_lc}
\end{listing}
%schriste - note to self, fix this example, add resampled points on top of plot?

@@

%Stuart: Where did the pandas section go?
%ARI - looks like it's up top

### 2014-01-28 3-3-HELIO.tex

The HELiophysics Integrated Observatory (\href{http://helio-vo.eu}{HELIO}) has 
put together a list of web services which allows scientists to query and 
discover data throughout the heliosphere (\textit{e.g.}, solar, planetary, 
inter-planetary data, \ldots)\citep{dps2012}.
HELIO has been built with a Service-Oriented Architecture, 
\textit{i.e.} its capabilities are split into a number of tasks that are 
implemented as separate services. 
\href{http://helio-vo.eu}{HELIO} comprises nine different public services, 
which allows scientiststo search in different catalogues of registered events, 
solar features, data from instruments in the heliosphere and other information 
like planets or spacecraft position in time. 
Additionally, \href{http://helio-vo.eu}{HELIO} provides a service that uses a 
propagation model to link the data in different points of the solar system by 
its original nature (\textit{e.g.}, Earth auroras are a signature of magnetic 
field disturbances produced few days before on the Sun).
In addition to the primary, web-based interface to 
\href{http://helio-vo.eu}{HELIO}, the services can be separately accessed.

SunPy's \texttt{hec} module provides an interface to the
\textit{HELIO Event Catalogue} (HEC) service, and this module was developed as
part of a Google Summer of Code (GSOC) project in 2013.
The HEC service currently provides access to 84 catalogues from different
sources (automated and manual classification, event lists from publications, \ldots).
As with all of the HELIO services, the HEC service provides results in VOTable 
data format (defined by IVOA \cite{ochsenbein_ivoa_2011}), and the \texttt{hec}
module parses this output using the \texttt{astropy.io.votable} package.
This format has the advantage of containing metadata with information like
data provenance and the performed query.

>>>>

The HELiophysics Integrated Observatory (\href{http://helio-vo.eu}{HELIO}) has 
put together a list of web services which allows scientists to query and 
discover data throughout the heliosphere, from solar, and magnetospheric to planetary and 
inter-planetary data \citep{dps2012}.
HELIO has been built with a Service-Oriented Architecture, 
\textit{i.e.} its capabilities are split into a number of tasks that are 
implemented as separate services. 
\href{http://helio-vo.eu}{HELIO} comprises nine different public services, 
which allows scientists to search different catalogues of registered events, 
solar features, data from instruments in the heliosphere and other information 
like planets or spacecraft position in time. 
Additionally, \href{http://helio-vo.eu}{HELIO} provides a service that uses a 
propagation model to link the data in different points of the solar system by 
its original nature (\textit{e.g.}, Earth auroras are a signature of magnetic 
field disturbances produced few days before on the Sun).
In addition to the primary, web-based interface to 
\href{http://helio-vo.eu}{HELIO}, the services can be separately accessed.

SunPy's \texttt{hec} module provides an interface to the
\textit{HELIO Event Catalogue} (HEC) service (not to be confused with the HEK, discussed in Section \ref{ssec:hek}), and this module was developed as
part of a Google Summer of Code (GSOC) project in 2013.
The HEC service currently provides access to 84 catalogues from different
sources.
As with all of the HELIO services, the HEC service provides results in VOTable 
data format (defined by IVOA \cite{ochsenbein_ivoa_2011}), and the \texttt{hec}
module parses this output using the \texttt{astropy.io.votable} package.
This format has the advantage of containing metadata with information like
data provenance and the performed query.

### 2014-01-28 3-5-Database.tex

The database can store and manage all data that can be read via SunPy's 
\texttt{io} subpackage, while direct integration with the \textsc{VSO} 
subpackage is supported.
It also possible to add entries manually by file or directory. The package also provides
a \texttt{fetch} method which provides a unified data search which includes both local files
and files on the \textsc{VSO} are searched. This reduces the likelihood of downloading the same file 
multiple times. When a file is added to the database the file is scanned and a file hash is produced. 
The current date is associated with the entry along with metadata summaries such 
as instrument, date of observation, field of view, etc. 
The database also provides the ability to associate custom meta-data to 
each database entry such as keywords, comments, and favouring, as well as 
querying the full metadata (\textit{i.e.} FITS header) of each entry.

>>>>

The database can store and manage all data that can be read via SunPy's 
\texttt{io} subpackage, while direct integration with the \textsc{VSO} 
subpackage is supported.
It is also possible to add entries manually by file or directory. The package also provides
a \texttt{fetch} method which provides a unified data search which includes both local files
and files on the \textsc{VSO}. This reduces the likelihood of downloading the same file 
multiple times. When a file is added to the database the file is scanned and a file hash is produced. 
The current date is associated with the entry along with metadata summaries such 
as instrument, date of observation, field of view, etc. 
The database also provides the ability to associate custom meta-data to 
each database entry such as keywords, comments, and favouring, as well as 
querying the full metadata (\textit{i.e.} FITS header) of each entry.

### 2014-01-28 4-Util.tex

\section{Additional functionality}\label{sec:util}
SunPy is meant to provide a consistent environment for solar data analysis. In 
order to achieve this goal SunPy provides a number of utility functions and packages which 
are used by the other SunPy modules and are made available to the user. This section 
briefly describes some of these functions.

>>>>

\section{Additional functionality}\label{sec:util}
SunPy is meant to provide a consistent environment for solar data analysis. In 
order to achieve this goal SunPy provides a number of additional functions and packages which 
are used by the other SunPy modules and are made available to the user. This section 
briefly describes some of these functions.

### 2014-01-28 6-Future.tex

As discussed in Section \ref{sec:Intro}, the primary focus of the 
\texttt{sunpy} library is the analysis and visualisation of `high-level' solar 
data. This means data that has been put through instrument processing 
and 
calibration routines, and contains full (WCS) coordinate information. 
The 
development plan for SunPy is to continue development within this 
scope. The 
primary components of this plan are to provide a set of data types 
that are 
interchangeable with one another: \textit{e.g.}, if you slice a 
\texttt{MapCube} 
along one coordinate, a \texttt{LightCurve} of intensity along the 
time range of 
the \texttt{MapCube} should be returned. To achieve this goal, all the 
data 
types need to share a unified coordinate system architecture so that 
each data 
type is aware of what the physical type of its data is and how 
operations on 
that data should be performed.

In concert with the work on the data types, further integration with 
the 
\texttt{astropy} package will enable SunPy to inherit the large 
amount of work 
done in that project to support a platform for astronomical data 
analysis. For 
example, the \texttt{astropy.units} submodule provides a simple and 
efficient 
way of representing physical units in Python data types; there are 
plans to 
incorporate this widely in the SunPy code base. In the past 12 months, 
the 
collaboration with the Astropy project 
\citep{theastropycollaboration2013} and 
its community has grown, abd this direct collaboration between Astropy 
% what abd means?
and 
SunPy will be one of the key strengths of SunPy going forward.

>>>>

As discussed in Section \ref{sec:Intro}, the primary focus of the 
SunPy library is the analysis and visualisation of `high-level' solar 
data. This means data that has been put through instrument processing 
and 
calibration routines, and contains full (WCS) coordinate information. 
The plan for SunPy is to continue development within this 
scope. The 
primary components of this plan are to provide a set of data types 
that are 
interchangeable with one another: \textit{e.g.}, if you slice a 
\texttt{MapCube} 
along one coordinate, a \texttt{LightCurve} of intensity along the 
time range of 
the \texttt{MapCube} should be returned. To achieve this goal, all the 
data 
types need to share a unified coordinate system architecture so that 
each data 
type is aware of what the physical type of its data is and how 
operations on 
that data should be performed.

In concert with the work on the data types, further integration with 
the 
\texttt{astropy} package will enable SunPy to inherit the large 
amount of work 
done in that project to support a platform for astronomical data 
analysis. For 
example, the \texttt{astropy.units} submodule provides a simple and 
efficient 
way of representing physical units in Python data types; there are 
plans to 
incorporate this widely in the SunPy code base. In the past 12 months, 
the 
collaboration with the Astropy project 
\citep{theastropycollaboration2013} and 
its community has grown, and this direct collaboration between Astropy 
% what abd means?
and 
SunPy will be one of the key strengths of SunPy going forward.
